---
title: "diamonds"
author: "Cynthia O'Donnell"
date: "05/29/2015"
output: html_document
---

# Analyzing Diamonds

```{r}
library(ggplot2)
data(diamonds)
summary(diamonds)
str(diamonds)
summary(diamonds$color)
```

#### A histogram of diamond prices

```{r, echo=FALSE}
qplot(x = price, data = diamonds, color=I('#040600'), fill=I('#8e0050'))
```
```{r}
summary(diamonds$price)
```

##### The distribution is right tailed with a median of 2401 and a mean of 3933.

```{r}
summary(diamonds$price < 500)
summary(diamonds$price < 250)
summary(diamonds$price > 15000)
```
There are no diamonds worth less than $250 and there are 1655 diamonds worth more than $15,000.

#### Let's change the bin width and the axes
```{r, echo=FALSE}
qplot(x=diamonds$price, data=diamonds, binwidth=50, color= I('#064e06')) +
  scale_x_continuous(limits=c(0,10000))
```

#### Facet Wrap by cut
```{r, echo=FALSE}
qplot(x=price, data = diamonds, color= I('#999999'), fill= I('#ffaa00')) +
 facet_wrap(~cut)
```
#### Which cut has the highest priced diamond?
```{r}
diamonds$cut[diamonds$price == max(diamonds$price)]
```
#### Which cut has the lowest priced diamond?
```{r}
diamonds$cut[diamonds$price == min(diamonds$price)]
```
#### Which cut has the lowest median price?
```{r}
by(diamonds$price, diamonds$cut, median)
```
#### Lets look into the relative price distributions by cut again.
```{r}
qplot(x = price, data = diamonds, color=I("#414141"), fill=I("#0641c8")) + 
  facet_wrap(~cut, scales="free_y")
```
##### Now the graphs indicate that the distributions are similarly distributed.

### Price Per Carat
```{r}
by(diamonds$price, diamonds$carat, summary)
qplot(x = price/carat, data = diamonds, color=I("#de007e"), fill=I("#414141"), binwidth = 10) + 
  facet_wrap(~cut, scales="free_y")
```


